---
title: "Is Louisville Crime on the Rise"
output: pdf_document
---

# Introduction
Over the past year I have gotten used to the constant flow of crime reports littering
the local news. Whether it was riots, teens running rampant or a record number of homicides,
it seemed Louisville's citizenry was spiraling out of control.  But the more I saw, 
the more I began to wonder--is crime really getting worse?  Is it no longer safe to walk
the streets? Or is there no difference at all?  Today, I aim to find out.

Luckily, the city of Louisville has a rich open crime dataset that is ripe for analysis.
Although records extend back to the 1920's, the data is very sparse up until around
2004.  For the purposes of this investigation, I will look at 2005-2015. 

# Overview of All Crime Incidents

My initial approach was just to look at overall crime distribution throughout the city.
As can be seen below, there is a definite hot spot in downtown and a lesser one near Shively
between I-264 and I-65. Of course, the next logical question, is has this base view changed
over the past 11 years.
```{r package_load, cache = TRUE, echo = FALSE, message=FALSE, warning = FALSE}
#Loading all packages necessary for report
library(stringi)
library(dplyr)
library(ggplot2)
library(ggmap)
library(readr)
library(printr)
library(viridis)
library(gridExtra)
```

```{r, load_data, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE}
# since most data cleaning was done outside this report, all I have to do here is
# create any useful additional variables and load the file
raw_data <- read_csv("crime_lou_with_geocoding.csv")

create_date_variables <- function(df){
  require(lubridate)
  # Uses the POSIXct date_occured variable to create useful date related
  # subvariables
  df$year <- year(df$date_occured)
  df$month <- month(df$date_occured)
  df$day <- day(df$date_occured)
  df$hour <- hour(df$date_occured)
  df$year_month <- paste(df$year, df$month, sep = '-')
  df$day_of_week <- wday(df$date_occured, label = TRUE, abbr = FALSE)
  df$weekday <- ifelse(df$day_of_week == "Saturday" | df$day_of_week == "Sunday", 
                              "Weekend", "Weekday")
  
  return(df)
}

crime_lou <- create_date_variables(raw_data)
```

```{r full_crime_plot, echo = FALSE, cache = TRUE, out.width = "1100px", warning = FALSE, message = FALSE}
density.theme <- base.theme <- 
  theme(axis.text = element_blank(), axis.ticks = element_blank())+
  theme(legend.title = element_text(size = 14, family = 'serif', face = 'bold', color = "#666666"))+
  theme(plot.title = element_text(size = 22, family = "serif", face = 'bold', hjust = 0, color = "#666666"))+
  theme(axis.title = element_blank())

bar.theme <- 
  theme(legend.position = "none")+
  theme(plot.title = element_text(size = 22, family = "serif", face = 'bold', hjust = 0, color = "#666666"))+
  theme(axis.title = element_text(size = 18, family = "serif", face = 'bold', hjust = 0, color = "#666666"))+
  theme(axis.title.y = element_text(angle = 90))+
  theme(axis.title.x = element_text(hjust = 0, vjust = 0.5))

  
crime_lou <- crime_lou%>%
  filter(year <=2015 & year >=2005 & !is.na(lat)  & !is.na(lng))

louisvilleMap <- get_map(location = c(lon = -85.686028, lat = 38.181602), source = "google",
                       maptype = 'roadmap', color = 'bw', zoom = 11)

ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = crime_lou,
    size = .00001, bins = 50, alpha = .8, color = "gray20"
  )+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  ggtitle("Louisville Crime, 2005-2015")+
  density.theme
```

As we can see here, the distribution of crimes has stayed roughly constant over the years.
```{r full_crime_yearly, echo = FALSE, cache = TRUE, out.width = '1100px', warning = FALSE, message = FALSE}
ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = crime_lou,
    size = .00001, bins = 50, alpha = .8, color = "gray20"
  )+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  ggtitle("Louisville Crime, By Year")+
  facet_wrap(~year)+
  density.theme
```

Looked at another way, the raw incidence reports appear to peak around 2009, but have since
leveled off.  While it does appear there is an upward trend, more analysis is needed before
drawing any conclusions.
```{r full_crime_bar, echo = FALSE, cache = TRUE, out.width = '1100px', warning = FALSE, message = FALSE}
crime_lou %>%
  group_by(year)%>%
  summarise(count = n())%>%
  ggplot(aes(x = year, y = count))+
  stat_smooth(method = 'lm', size = 1.25, alpha = .8, fill = "grey", colour = "black")+
  geom_bar(stat = 'identity', colour = "black",fill = "darkblue",  alpha = .7)+
  ggtitle("Incident Count by Year")+
  labs(x = "Year", y = "Incident Count")+
  scale_x_discrete(limits = seq(2005, 2015, by = 1))+
  bar.theme
```


# Violent Crime

In order to really see how crime in the city has changed, I need to break it down by
crime type. With all the talk of record homicide numbers in 2015, I decided to start 
with violent crime. There are four offenses the FBI considers violent crime -- murder and 
nonnegligent manslaughter, forcible rape, robbery and aggravated assault.  Looking at the 
aggregate, a minor increase is apparent.

```{r violent_crime_full, echo = FALSE, cache = TRUE, out.width = '1100px', warning = FALSE, message = FALSE}
crime_lou%>%
  filter(nibrs_code == '13a' |nibrs_code == '120'| nibrs_code == '09a'|nibrs_code == '11a')%>%
  group_by(year)%>%
  summarise(count = n())%>%
  ggplot(aes(x = year, y = count))+
  stat_smooth(method = 'lm', size = 1.25, alpha = .8, fill = "grey", colour = "black")+
  geom_bar(stat = 'identity', colour = "black",fill = "darkblue",  alpha = .7)+
  ggtitle("Violent Crime By Year")+
  labs(x = "Year", y = "Incident Count")+
  scale_x_discrete(limits = seq(2005, 2015, by = 1))+
  bar.theme

```

But a more fine grained view shows that the story is a little more complicated.  While 
homicide and assaults are indeed up, rape and robberies are down.
```{r violent_crime_by_year, echo = FALSE, cache = TRUE, out.width = '1100px', warning = FALSE, message = FALSE}
crime_lou%>%
  filter(nibrs_code == '09a'|nibrs_code == '11a')%>%
  group_by(year, crime_type)%>%
  summarise(count = n())%>%
  ggplot(aes(x = year, y = count))+
  stat_smooth(method = 'lm', size = 1.25, alpha = .8, fill = "grey", colour = "black")+
  geom_bar(stat = 'identity', colour = "black",fill = "darkblue",  alpha = .7)+
  facet_wrap(~crime_type)+
  ggtitle("Homicide and Rape Incidents")+
  labs(x = "Year", y = "Incident Count")+
  scale_x_discrete(limits = seq(2005, 2015, by = 1))+
  bar.theme

crime_lou%>%
  filter(nibrs_code == '13a'|nibrs_code == '120')%>%
  group_by(year, crime_type)%>%
  summarise(count = n())%>%
  ggplot(aes(x = year, y = count))+
  stat_smooth(method = 'lm', size = 1.25, alpha = .8, fill = "grey", colour = "black")+
  geom_bar(stat = 'identity', colour = "black",fill = "darkblue",  alpha = .7)+
  facet_wrap(~crime_type)+
  ggtitle("Aggravated Assault and Robbery Incidents")+
  labs(x = "Year", y = "Incident Count")+
  scale_x_discrete(limits = seq(2005, 2015, by = 1))+
  bar.theme
```

```{r, violent_crime_percentage_changes, cache = TRUE, echo = FALSE}
violent_sum_df <- crime_lou%>%
  filter(nibrs_code == '13a' |nibrs_code == '120'| nibrs_code == '09a'|nibrs_code == '11a')%>%
  group_by(year, crime_type)%>%
  summarise(count = n())


vio_per_changes <- violent_sum_df%>%
  group_by(crime_type)%>%
  mutate(Change_from_2014 = paste0(round(100*((count /lag(count))-1), 1), "%"),
         Change_from_Max = paste0(round(100*((count /max(count))-1), 1), "%"),
         Change_from_Median = paste0(round(100*((count /median(count))-1), 1), "%"))%>%
  filter(year == 2015)%>%
  select(crime_type, Change_from_2014, Change_from_Max, Change_from_Median)

all_vio_sum <- crime_lou%>%
  filter(nibrs_code == '13a' |nibrs_code == '120'| nibrs_code == '09a'|nibrs_code == '11a')%>%
  group_by(year)%>%
  summarise(count = n())


all_vio_changes <- all_vio_sum%>%
  mutate(crime_type = 'all',
         Change_from_2014 = paste0(round(100*((count /lag(count))-1), 1), "%"),
         Change_from_Max = paste0(round(100*((count /max(count))-1), 1), "%"),
         Change_from_Median = paste0(round(100*((count /median(count))-1), 1), "%"))%>%
  filter(year == 2015)%>%
  select(crime_type, Change_from_2014, Change_from_Max, Change_from_Median)
  
all_vio_changes <- rbind(all_vio_changes, vio_per_changes)

all_vio_changes
```


Looking at the distribution of the violent crimes, you see the family hot spots in downtown
and near Shively.

```{r violent_crime_distribution, echo = FALSE, cache = TRUE, out.width = '1100px', warning = FALSE, message = FALSE}
mapData <- crime_lou%>%
  filter(nibrs_code == '13a' |nibrs_code == '120'| nibrs_code == '09a'|nibrs_code == '11a'& 
           !is.na(lat)  & !is.na(lng))
  
ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = mapData,
    size = .00001, bins = 75, alpha = .8, color = "gray20"
  )+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  ggtitle("Violent Crime, 2005-2015")+
  density.theme

density.theme <- base.theme <- 
  theme(axis.text = element_blank(), axis.ticks = element_blank())+
  theme(legend.title = element_text(size = 14, family = 'serif', face = 'bold', color = "#666666"))+
  theme(plot.title = element_text(size = 18, family = "serif", face = 'bold', hjust = 0, color = "#666666"))+
  theme(axis.title = element_blank())

p1 <- ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = mapData%>%filter(nibrs_code == '13a'),
    size = .00001, bins = 40, alpha = .8, color = "gray20"
  )+
  ggtitle("Distribution of Assaults")+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  density.theme

p2 <- ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = mapData%>%filter(nibrs_code == '120'),
    size = .00001, bins = 40, alpha = .8, color = "gray20"
  )+
  ggtitle("Distribution of Robberies")+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  density.theme

p3 <- ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = mapData%>%filter(nibrs_code == '09a'),
    size = .00001, bins = 40, alpha = .8, color = "gray20"
  )+
  ggtitle("Distribution of Homicides")+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  density.theme

p4 <- ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = mapData%>%filter(nibrs_code == '11a'),
    size = .00001, bins = 40, alpha = .8, color = "gray20"
  )+
  ggtitle("Distribution of Rapes")+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  density.theme

grid.arrange(p1, p2, p3, p4, ncol = 2)
```

# Theft Crimes
What about if we focus solely on theft related offenses?  Do we see the same sort of 
distribution and trend, or have these property crimes become more prevalent or 
more widely dispersed?


Again, I start by looking at the aggregate of all theft related crimes.  This includes 
burglary, motor vehicle theft, and 8 other subcategories of theft/larceny.  As you can see,
there is a slight upward trend, but overall, 2015 appears to be a typical year in Louisville
vis-a-vis thefts.
```{r theft_crime_full, echo = FALSE, cache = TRUE, out.width = '1100px', warning = FALSE, message = FALSE}
crime_lou%>%
  # filter(nibrs_code == '240'|nibrs_code == '23a'|nibrs_code =='23b'|nibrs_code == '23c'|
  #          nibrs_code == '23d'|nibrs_code =='23e'|nibrs_code == '23f'|nibrs_code == '23g'|
  #          nibrs_code == '23h'|nibrs_code == '220')
  filter(crime_type == 'motor vehicle theft'|crime_type == 'theft/larceny'|crime_type == 'burglary'|
           crime_type == 'vehicle break-in/theft')%>%
  group_by(year)%>%
  summarise(count = n())%>%
  ggplot(aes(x = year, y = count))+
  stat_smooth(method = 'lm', size = 1.25, alpha = .8, fill = "grey", colour = "black")+
  geom_bar(stat = 'identity', colour = "black",fill = "darkblue",  alpha = .7)+
  ggtitle("Thefts By Year")+
  labs(x = "Year", y = "Incident Count")+
  scale_x_discrete(limits = seq(2005, 2015, by = 1))+
  bar.theme
```

A more detailed breakdown of thefts shows us that most categories are on the decline, 
but general theft/larceny(which is roughly the same size as the other three categories combined)
seems to show a significant upward trend.
```{r theft_crime_by_year, echo = FALSE, cache = TRUE, out.width = '1100px', warning = FALSE, message = FALSE}
crime_lou%>%
  filter(crime_type == 'motor vehicle theft'|crime_type == 'theft/larceny'|crime_type == 'burglary'|
           crime_type == 'vehicle break-in/theft')%>%
  group_by(year, crime_type)%>%
  summarise(count = n())%>%
  ggplot(aes(x = year, y = count))+
  stat_smooth(method = 'lm', size = 1.25, alpha = .8, fill = "grey", colour = "black")+
  geom_bar(stat = 'identity', colour = "black",fill = "darkblue",  alpha = .7)+
  facet_wrap(~crime_type)+
  ggtitle("Thefts By Year")+
  labs(x = "Year", y = "Incident Count")+
  scale_x_discrete(limits = seq(2005, 2015, by = 1))+
  bar.theme
```

```{r, theft_crime_percentage_changes, cache = TRUE, echo = FALSE}
theft_sum_df <- crime_lou %>%
  filter(crime_type == "burglary"|crime_type == 'motor vehicle theft'|crime_type == 'theft/larceny'|
           crime_type == 'vehicle break-in/theft')%>%
  group_by(year, crime_type)%>%
  summarise(count = n())

theft_per_changes <- theft_sum_df%>%
  group_by(crime_type)%>%
  mutate(Change_from_2014 = paste0(round(100*((count /lag(count))-1), 1), "%"),
         Change_from_Max = paste0(round(100*((count /max(count))-1), 1), "%"),
         Change_from_Median = paste0(round(100*((count /median(count))-1), 1), "%"))%>%
  filter(year == 2015)%>%
  select(crime_type, Change_from_2014, Change_from_Max, Change_from_Median)

all_theft_sum <- crime_lou%>%
  filter(crime_type == "burglary"|crime_type == 'motor vehicle theft'|crime_type == 'theft/larceny'|
           crime_type == 'vehicle break-in/theft')%>%
  group_by(year)%>%
  summarise(count = n())

all_theft_changes <- all_theft_sum%>%
  mutate(crime_type = 'all',
         Change_from_2014 = paste0(round(100*((count /lag(count))-1), 1), "%"),
         Change_from_Max = paste0(round(100*((count /max(count))-1), 1), "%"),
         Change_from_Median = paste0(round(100*((count /median(count))-1), 1), "%"))%>%
  filter(year == 2015)%>%
  select(crime_type, Change_from_2014, Change_from_Max, Change_from_Median)
  
all_theft_changes <- rbind(all_theft_changes, theft_per_changes)
all_theft_changes
```

```{r theft_crime_distribution, echo = FALSE, cache = TRUE, out.width = '1100px', warning = FALSE, message = FALSE}
mapData <- crime_lou%>%
  filter((crime_type == "burglary"|crime_type == 'motor vehicle theft'|crime_type == 'theft/larceny'|
           crime_type == 'vehicle break-in/theft') &  !is.na(lat)  & !is.na(lng))
  
ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = mapData,
    size = .00001, bins = 75, alpha = .8, color = "gray20"
  )+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  ggtitle("Louisville Thefts, 2005-2015")+
  density.theme

density.theme <- base.theme <- 
  theme(axis.text = element_blank(), axis.ticks = element_blank())+
  theme(legend.title = element_text(size = 14, family = 'serif', face = 'bold', color = "#666666"))+
  theme(plot.title = element_text(size = 18, family = "serif", face = 'bold', hjust = 0, color = "#666666"))+
  theme(axis.title = element_blank())

p1 <- ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = mapData%>%filter(crime_type == 'burglary'),
    size = .00001, bins = 40, alpha = .8, color = "gray20"
  )+
  ggtitle("Distribution of Burglary's")+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  density.theme

p2 <- ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = mapData%>%filter(crime_type == 'motor vehicle theft'),
    size = .00001, bins = 40, alpha = .8, color = "gray20"
  )+
  ggtitle("Distribution of Motor Vehicle Thefts")+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  density.theme

p3 <- ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = mapData%>%filter(crime_type == 'theft/larceny'),
    size = .00001, bins = 40, alpha = .8, color = "gray20"
  )+
  ggtitle("Distribution of theft/larceny")+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  density.theme

p4 <- ggmap(louisvilleMap)+
  stat_bin2d(
    aes(x = lng, y = lat),
    data = mapData%>%filter(crime_type == 'vehicle break-in/theft'),
    size = .00001, bins = 40, alpha = .8, color = "gray20"
  )+
  ggtitle("Distribution of Vehicle Break-in/theft")+
  scale_fill_viridis(option = "inferno", name = "Incident Count")+
  density.theme

grid.arrange(p1, p2, p3, p4, ncol = 2)
```